Simulations to test ctwas summary stats version, 113k samples, a few new setings required

### 2023-05-08 analysis/Paper_figures_simulation.Rmd
110k simulation

outputdir = "/home/simingz/causalTWAS/simulations/simulation_ctwas_rss_20210416/"
comparedir = "/home/simingz/causalTWAS/simulations/simulation_ctwas_rss_20210416_compare/"
runtag = "ukb-s80.45-adi"
configtag = 1
pgenfn = "/home/simingz/causalTWAS/ukbiobank/ukb_pgen_s80.45/ukb-s80.45_pgenfs.txt"
ld_pgenfn = "/home/simingz/causalTWAS/ukbiobank/ukb_pgen_s80.45/ukb-s80.45.2_pgenfs.txt"
exprfn = "/home/simingz/causalTWAS/simulations/simulation_ctwas_rss_20210416//ukb-s80.45-adi.expr.txt"

@@

@@

```{r pipplot, fig.width= 8, fig.height= 16}
simutaglist = lapply(c(1:3,5:9), function(x) paste(x, 1:5, sep ="-"))
mtxlist <- get_params(configtag, runtag, simutaglist)
simutaglist = c(list(paste(1, c(1,2,5), sep = "-")), 
                lapply(c(2,3,5,6,7,8,9), function(x) paste(x, 1:5, sep ="-")))

@@

### Comparison with other methods
Bar plot: each bar shows the number of genes, colored by causal status. Use a different color for each method. The method and cut off values:
* ctwas: PIP 0.8
* FUSION fdr: 0.05
* FUSION bonferroni: 0.05
* COLOC PP4: 0.8
* FOCUS PIP: 0.8
* SMR FDR: 0.05
* SMR HEIDI: HEIDI p > 0.05, SMR FDR < 0.05

@@

get_ncausal_df <- function(pfiles, cau, cut = 0.8, 
                           method = c("ctwas", "fusionfdr", "fusionbon","coloc", "focus", "smr", "smrheidi")){
  df <- NULL
  for (i in 1:length(pfiles)) {
    res <- fread(pfiles[i], header = T)
    # res <- res[complete.cases(res),]
    if (method == "ctwas"){
        res <- data.frame(res[res$type  =="gene", ])
        res$ifcausal <- ifelse(res$id %in% cau[[i]], 1, 0)
        res <- res[res$susie_pip > cut,]
    } else if (method == "fusionfdr"){
        res$FDR <- p.adjust(res$TWAS.P, method = "fdr")
        res <- res[res$FDR < cut,]
        res$ifcausal <- ifelse(res$ID %in% cau[[i]], 1, 0)
    } else if  (method == "fusionbon"){
        res$FDR <- p.adjust(res$TWAS.P, method = "bonferroni")
        res <- res[res$FDR < cut,]
        res$ifcausal <- ifelse(res$ID %in% cau[[i]], 1, 0)
    } else if  (method == "coloc"){
        res <- res[res$COLOC.PP4 > cut,]
        res$ifcausal <- ifelse(res$ID %in% cau[[i]], 1, 0)
    } else if (method == "focus"){
        res <- res[res$mol_name != "NULL",]
        res <- res[res$pip > cut, ] 
        res$ifcausal <- ifelse(res$mol_name %in% cau[[i]], 1, 0)
    } else if (method == "smr"){
        res$FDR <- p.adjust(res$p_SMR, method = "fdr")
        res <- res[res$FDR < cut,]
        res$ifcausal <- ifelse(res$probeID %in% cau[[i]], 1, 0)
    } else if (method == "smrheidi"){
        res <- res[res$p_HEIDI > 0.05,]
        res$FDR <- p.adjust(res$p_SMR, method = "fdr")
        res <- res[res$FDR < cut,]
        res$ifcausal <- ifelse(res$probeID %in% cau[[i]], 1, 0)
    } else{
      stop("no such method")
    }

@@

  susieIfs <- paste0(outputdir, runtag, "_simu",simutags, "_config", configtag,".susieIrss.txt")
  fusioncolocfs <- paste0(comparedir, runtag, "_simu", simutags, ".Adipose_Subcutaneous.coloc.result")
  focusfs <- paste0(comparedir, runtag, "_simu", simutags, ".Adipose_Subcutaneous.focus.tsv")
  smrfs <- paste0(comparedir, runtag, "_simu", simutags, ".Adipose_Subcutaneous.smr")

  ctwas_df <- get_ncausal_df(susieIfs, cau= cau, cut = 0.8, method ="ctwas")
  fusionfdr_df <- get_ncausal_df(fusioncolocfs, cau= cau, cut = 0.05, method = "fusionfdr")
  fusionbon_df <- get_ncausal_df(fusioncolocfs , cau= cau, cut = 0.05, method = "fusionbon")
  coloc_df <- get_ncausal_df(fusioncolocfs , cau= cau, cut = 0.8, method = "coloc")
  focus_df <- get_ncausal_df(focusfs , cau= cau, cut = 0.8, method = "focus")
  smr_df <- get_ncausal_df(smrfs, cau= cau, cut = 0.05, method = "smr")
  smrheidi_df <- get_ncausal_df(smrfs, cau= cau, cut = 0.05, method = "smrheidi")

  df <- rbind(ctwas_df, fusionfdr_df, fusionbon_df, coloc_df, focus_df, smr_df, smrheidi_df)
  df$ifcausal <- df$ifcausal + as.numeric(as.factor(df$method))*10
  df$ifcausal <- as.factor(df$ifcausal)
  fig <- ggbarplot(df, x = "method", y = "count", add = "mean_se", fill = "ifcausal", palette = colors, legend = "none", ...) + grids(linetype = "dashed")
  fig
}

```{r ncausalplot, fig.width= 10, fig.height= 4}
colset = c("#ebebeb", "#bebada" , "#ebebeb", "#fb8072", "#ebebeb","#ffffb3","#ebebeb", "#8dd3c7", "#ebebeb","#8dd3c7", "#ebebeb", "#87CEFA", "#ebebeb","#87CEFA") # coloc, ctwas, focus, fusionbon, fusionfdr, smrheidi, smr
f1 <- plot_ncausal(configtag, runtag,  simutaglist[[1]], colors = colset, ylim= c(0,150), main = "high gene PVE")
f2 <- plot_ncausal(configtag, runtag,  simutaglist[[2]], colors = colset, ylim= c(0,150), main = "low gene PVE")
gridExtra::grid.arrange(f1, f2, ncol=2)
```

@@

  if (isTRUE(rerun_ctwas)){
      ld_exprfs <- paste0(outputdir, runtag, "_simu", simutag, "_chr", 1:22, ".expr.gz")
      load(paste0(outputdir, runtag, "_simu", simutag, "_config", configtag, ".s2.susieIrssres.Rd"))
      source(paste0(outputdir, "config", configtag, ".R"))
      group_prior <- group_prior_rec[, ncol(group_prior_rec)] 
      group_prior[2] <- group_prior[2] * thin
      group_prior_var <- group_prior_var_rec[, ncol(group_prior_var_rec)]
      temp_reg <- data.frame("chr" = paste0("chr",chr), "start" = startpos, "stop" = endpos)
      write.table(temp_reg, file= "temp_reg.txt" , row.names=F, col.names=T, sep="\t", quote = F)
      z_gene <-  a[a$type == "gene", c("id", "t.value")]
      colnames(z_gene) <- c("id", "z")
      z_snp <-  a[a$type == "SNP", c("id", "alt", "ref", "t.value")]
      colnames(z_snp) <- c("id", "A1", "A2", "z")
      ctwas_rss(z_gene, z_snp, ld_exprfs= ld_exprfs, ld_pgenfs = ld_pgenfs, ld_R_dir = NULL, ld_regions_custom = "temp_reg.txt", thin = 1, outputdir = ".", outname = "temp", ncore = 1, ncore.rerun = 1, prob_single = 0,  group_prior = group_prior, group_prior_var = group_prior_var, estimate_group_prior = F, estimate_group_prior_var = F, recover_strand_ambig_z = F)
      a2 <- fread("temp.susieIrss.txt", header = T)
      a <- merge(a2, b, by = "id", all = T)
      # file.remove("temp.susieIrss.txt")
      # file.remove("temp.temp.susieIrssres.Rd")
      # file.remove("temp.regions.txt")
      # file.remove("temp_reg.txt")
   }

@@

### coloc false positives
```{r regionplot4, echo=FALSE,results='hide',fig.keep='all', message=F}
simutag <- "3-4"
a <- plot_region(runtag, simutag, configtag, chr = 1, startpos = 64000000 , endpos =66000000, rerun_ctwas = F, plot_eqtl = T )
```
```{r regionplot4, echo=FALSE,results='hide',fig.keep='all', message=F}
simutag <- "3-4"
a <- plot_region(runtag, simutag, configtag, chr = 1, startpos = 66000000 , endpos =70000000, rerun_ctwas = F, plot_eqtl = T )
```
```{r regionplot4, echo=FALSE,results='hide',fig.keep='all', message=F}
simutag <- "3-4"
a <- plot_region(runtag, simutag, configtag, chr = 1, startpos = 155000000 , endpos =170000000, rerun_ctwas = F, plot_eqtl = T )
```
```{r regionplot4, echo=FALSE,results='hide',fig.keep='all', message=F}
simutag <- "3-4"
a <- plot_region(runtag, simutag, configtag, chr = 5, startpos = 47000000 , endpos =50000000, rerun_ctwas = F, plot_eqtl = T )
```

>>>>

outputdir = "/home/simingz/causalTWAS/simulations/simulation_ctwas_rss_20210416/"
comparedir = "/home/simingz/causalTWAS/simulations/simulation_ctwas_rss_20210416_compare/"
runtag = "ukb-s80.45-adi"
configtag = 1
pgenfn = "/home/simingz/causalTWAS/ukbiobank/ukb_pgen_s80.45/ukb-s80.45_pgenfs.txt"
ld_pgenfn = "/home/simingz/causalTWAS/ukbiobank/ukb_pgen_s80.45/ukb-s80.45.2_pgenfs.txt"
exprfn = "/home/simingz/causalTWAS/simulations/simulation_ctwas_rss_20210416//ukb-s80.45-adi.expr.txt"
weightf = "/project2/mstephens/causalTWAS/fusion_weights/Adipose_Subcutaneous.pos"

@@

weights <- as.data.frame(fread(weightf, header = T))
weights$ENSEMBL_ID <- sapply(weights$WGT, function(x){unlist(strsplit(unlist(strsplit(x,"/"))[2], "[.]"))[2]})

@@

```{r pipplot-suppl, fig.width= 8, fig.height= 16, eval=F}
simutaglist = lapply(c(1:3,5:9), function(x) paste(x, 1:5, sep ="-"))
mtxlist <- get_params(configtag, runtag, simutaglist)
simutaglist = c(list(paste(1, c(1,2,5), sep = "-")), 
                lapply(c(2,3,5,6,7,8,9), function(x) paste(x, 1:5, sep ="-")))

@@

### Comparison with other methods
Bar plot: each bar shows the number of genes, colored by causal status. Use a different color for each method. The method and cut off values:
* ctwas: PIP 0.8
* FUSION fdr: 0.05
* FUSION bonferroni: 0.05
* FUSION permutation fdr: 0.05
* COLOC PP4: 0.8
* FOCUS PIP: 0.8
* SMR FDR: 0.05, updated using ensemble ID
* SMR HEIDI: HEIDI p > 0.05, SMR FDR < 0.05, updated using ensemble ID

@@

get_ncausal_df <- function(pfiles, cau, cut = 0.8, useFDP= F,
                           method = c("ctwas", "fusionfdr", "fusionbon","fusionperm", "coloc", "focus", "smr", "smrheidi")){
  df <- NULL
  for (i in 1:length(pfiles)) {
    res <- fread(pfiles[i], header = T)
    # res <- res[complete.cases(res),]
    if (method == "ctwas"){
        res <- data.frame(res[res$type  =="gene", ])
        res$ifcausal <- ifelse(res$id %in% cau[[i]], 1, 0)
        res <- res[order(res$susie_pip, decreasing = T ),] 
        res$FDP <- 1-cumsum(res$ifcausal)/(1:nrow(res))
        if (isTRUE(useFDP)){ res <- res[res$FDP < cut,]} else {
          res <- res[res$susie_pip > cut,]}
    } else if (method == "fusionfdr"){
        res$FDR <- p.adjust(res$TWAS.P, method = "fdr")
        res <- res[order(res$FDR, decreasing = F),] 
        res$ifcausal <- ifelse(res$ID %in% cau[[i]], 1, 0)
        res$FDP <- 1-cumsum(res$ifcausal)/(1:nrow(res))
        if (isTRUE(useFDP)){ res <- res[res$FDP < cut,]} else {
          res <- res[res$FDR < cut,]}

    } else if  (method == "fusionbon"){
        res$FDR <- p.adjust(res$TWAS.P, method = "bonferroni")
        res <- res[order(res$FDR, decreasing = F),] 
        res$ifcausal <- ifelse(res$ID %in% cau[[i]], 1, 0)
        res$FDP <- 1-cumsum(res$ifcausal)/(1:nrow(res))
        if (isTRUE(useFDP)){ res <- res[res$FDP < cut,]} else {
          res <- res[res$FDR < cut,]
        }
    } else if  (method == "fusionperm"){
        res <- res[res$PERM.N!=0,]
        res$FDR <- p.adjust(res$PERM.PV, method = "fdr")
        res <- res[order(res$FDR, decreasing = F),] 
        res$ifcausal <- ifelse(res$ID %in% cau[[i]], 1, 0)
        res$FDP <- 1-cumsum(res$ifcausal)/(1:nrow(res))
        if (isTRUE(useFDP)){ res <- res[res$FDP < cut,]} else {
          res <- res[res$FDR < cut,]
        }
    } else if  (method == "coloc"){
        res <- res[!is.na(res$COLOC.PP4),]
        res <- res[order(res$COLOC.PP4, decreasing = T),] 
        res$ifcausal <- ifelse(res$ID %in% cau[[i]], 1, 0)
        res$FDP <- 1-cumsum(res$ifcausal)/(1:nrow(res))
        if (isTRUE(useFDP)){ res <- res[res$FDP < cut,]} else {
         res <- res[res$COLOC.PP4 > cut,]
        }
    } else if (method == "focus"){
        res <- res[res$mol_name != "NULL",]
        res$ifcausal <- ifelse(res$mol_name %in% cau[[i]], 1, 0)
        res <- res[order(res$pip, decreasing = T),] 
        res$FDP <- 1-cumsum(res$ifcausal)/(1:nrow(res))
        if (isTRUE(useFDP)){ res <- res[res$FDP < cut,]} else {
          res <- res[res$pip > cut, ] 
        }
    }  else if (method == "SMR"){
        res <- as.data.frame(res)
        res$probeID <- sapply(res$Gene, function(x){unlist(strsplit(x, "[.]"))[1]})
        res <- res[res$probeID %in% weights$ENSEMBL_ID,]
        res <- res[sapply(res$p_HEIDI > 0.05, isTRUE),]
        res$FDR <- p.adjust(res$p_SMR, method = "fdr")
        res$ifcausal <- ifelse(res$probeID %in% cau[[i]], 1, 0)
        res <- res[order(res$FDR, decreasing = F),] 
        res$FDP <- 1-cumsum(res$ifcausal)/(1:nrow(res))
        if (isTRUE(useFDP)){ res <- res[res$FDP < cut,]} else {
           res <- res[res$FDR < cut,]
        }
    } else if (method == "MR-JTI"){
        res$ifcausal <- ifelse(res$variable %in% cau[[i]], 1, 0)
        res <- res[order(abs(res$beta), decreasing = T),] 
        res$FDP <- 1-cumsum(res$ifcausal)/(1:nrow(res))
        if (isTRUE(useFDP)){ res <- res[res$FDP < cut,]} else {
          res <- res[res$CI_significance=="sig",] 
        }
    } else if (method == "PMR-Egger"){
        res <- as.data.frame(res)
        res$ifcausal <- ifelse(res$gene_id %in% cau[[i]], 1, 0) 
        res <- res[!is.na(res$causal_pvalue), ]
        res <- res[order(res$causal_pvalue, decreasing = F),] 
        res$FDP <- 1-cumsum(res$ifcausal)/(1:nrow(res))
        if (isTRUE(useFDP)){ res <- res[res$FDP < cut,]} else {
          res <- res[sapply(res$causal_pvalue < cut/sum(!is.na(res$causal_pvalue)), isTRUE),]
        }
    } else{
      stop("no such method")
    }

@@

  cau_ensembl <- cau

  for (i in 1:length(cau_ensembl)){
    cau_ensembl[[i]][cau_ensembl[[i]] %in% weights$ID] <- weights$ENSEMBL_ID[match(cau_ensembl[[i]][cau_ensembl[[i]] %in% weights$ID], weights$ID)]
  }

  susieIfs <- paste0(outputdir, runtag, "_simu",simutags, "_config", configtag,".susieIrss.txt")
  fusioncolocfs <- paste0(comparedir, runtag, "_simu", simutags, ".Adipose_Subcutaneous.coloc.result")
  focusfs <- paste0(comparedir, runtag, "_simu", simutags, ".Adipose_Subcutaneous.focus.tsv")
  smrfs <- paste0(comparedir, runtag, "_simu", simutags, ".Adipose_Subcutaneous.smr")
  mrjtifs <- paste0(comparedir, runtag, "_simu", simutags, ".Adipose_Subcutaneous.mrjti.result")
  pmrfs <- paste0(comparedir, runtag, "_simu", simutags, ".Adipose_Subcutaneous.pmr.result_pi_080")

  ctwas_df <- get_ncausal_df(susieIfs, cau= cau, cut = 0.8, method ="ctwas")
  #fusionfdr_df <- get_ncausal_df(fusioncolocfs, cau= cau, cut = 0.05, method = "fusionfdr")
  fusionbon_df <- get_ncausal_df(fusioncolocfs , cau= cau, cut = 0.05, method = "fusionbon")
  fusionperm_df <- get_ncausal_df(fusioncolocfs , cau= cau, cut = 0.05, method = "fusionperm",)
  coloc_df <- get_ncausal_df(fusioncolocfs , cau= cau, cut = 0.8, method = "coloc")
  focus_df <- get_ncausal_df(focusfs , cau= cau, cut = 0.8, method = "focus")
  #smr_df <- get_ncausal_df(smrfs, cau= cau, cut = 0.05, method = "smr")
  smrheidi_df <- get_ncausal_df(smrfs, cau= cau_ensembl, cut = 0.05, method = "SMR")
  mrjti_df <- get_ncausal_df(mrjtifs, cau= cau_ensembl, method = "MR-JTI")
  pmr_df <- get_ncausal_df(pmrfs, cau= cau_ensembl, cut = 0.05, method = "PMR-Egger")

  df <- rbind(ctwas_df, fusionbon_df,fusionperm_df, coloc_df, focus_df, smrheidi_df, mrjti_df, pmr_df)
  df$ifcausal <- df$ifcausal + as.numeric(as.factor(df$method))*10
  df$ifcausal <- as.factor(df$ifcausal)
  fig <- ggbarplot(df, x = "method", y = "count", add = "mean_se", fill = "ifcausal", palette = colors, legend = "none", ...) + grids(linetype = "dashed")
  fig
}
```

```{r ncausalplot, fig.width= 10, fig.height= 4}
colset = c("#ebebeb", "#ffffb3", # FOCUS
           "#ebebeb", "#8dd3c7", # Fusion
           "#ebebeb", "palegreen", # Fusion-permutation
           "#ebebeb", "#CC79A7", # MR-JTI
           "#ebebeb", "goldenrod", #PMR-Egger
           "#ebebeb", "#87CEFA", # SMR
           "#ebebeb", "#fb8072", # cTWAS
           "#ebebeb", "#bebada") # coloc
f1 <- plot_ncausal(configtag, runtag,  simutaglist[[1]], colors = colset, ylim= c(0,225), main = "high gene PVE")
f2 <- plot_ncausal(configtag, runtag,  simutaglist[[2]], colors = colset, ylim= c(0,225), main = "low gene PVE")
gridExtra::grid.arrange(f1, f2, ncol=2)
```

### Comparison with other methods at specified FDP
We can compare the power of methods at a given false discovery proportion, say 10% or 25%. If a method cannot achieve that FDP no matter what the threshold is used, we will point out that fact.

```{r ncausalplot-FDP, fig.width= 10, fig.height= 4}

plot_ncausal_FDP <- function(configtag, runtag,  simutags, colors, cut =0.2, ...){
  phenofs <- paste0(outputdir, runtag, "_simu", simutags, "-pheno.Rd")
  cau <- lapply(phenofs, function(x) {load(x);get_causal_id(phenores)})

  cau_ensembl <- cau

  for (i in 1:length(cau_ensembl)){
    cau_ensembl[[i]][cau_ensembl[[i]] %in% weights$ID] <- weights$ENSEMBL_ID[match(cau_ensembl[[i]][cau_ensembl[[i]] %in% weights$ID], weights$ID)]
  }

  susieIfs <- paste0(outputdir, runtag, "_simu",simutags, "_config", configtag,".susieIrss.txt")
  fusioncolocfs <- paste0(comparedir, runtag, "_simu", simutags, ".Adipose_Subcutaneous.coloc.result")
  focusfs <- paste0(comparedir, runtag, "_simu", simutags, ".Adipose_Subcutaneous.focus.tsv")
  smrfs <- paste0(comparedir, runtag, "_simu", simutags, ".Adipose_Subcutaneous.smr")
  mrjtifs <- paste0(comparedir, runtag, "_simu", simutags, ".Adipose_Subcutaneous.mrjti.result")
  pmrfs <- paste0(comparedir, runtag, "_simu", simutags, ".Adipose_Subcutaneous.pmr.result_pi_080")

  ctwas_df <- get_ncausal_df(susieIfs, cau= cau, cut = cut, method ="ctwas", useFDP = T)
  #fusionfdr_df <- get_ncausal_df(fusioncolocfs, cau= cau, cut = 0.05, method = "fusionfdr")
  fusionbon_df <- get_ncausal_df(fusioncolocfs , cau= cau, cut = cut, method = "fusionbon", useFDP = T)
  fusionperm_df <- get_ncausal_df(fusioncolocfs , cau= cau, cut = cut, method = "fusionperm", useFDP = T)
  coloc_df <- get_ncausal_df(fusioncolocfs , cau= cau, cut = cut, method = "coloc", useFDP = T)
  focus_df <- get_ncausal_df(focusfs , cau= cau, cut = cut, method = "focus", useFDP = T)
  #smr_df <- get_ncausal_df(smrfs, cau= cau, cut = 0.05, method = "smr")
  smrheidi_df <- get_ncausal_df(smrfs, cau= cau_ensembl, cut = cut, method = "SMR", useFDP = T)
  mrjti_df <- get_ncausal_df(mrjtifs, cau= cau_ensembl, cut = cut, method = "MR-JTI", useFDP = T)
  pmr_df <- get_ncausal_df(pmrfs, cau= cau_ensembl, cut = cut, method = "PMR-Egger", useFDP = T)

  df <- rbind(ctwas_df, fusionbon_df,fusionperm_df, coloc_df, focus_df, smrheidi_df, mrjti_df, pmr_df)
  df$ifcausal <- df$ifcausal + as.numeric(as.factor(df$method))*10
  df$ifcausal <- as.factor(df$ifcausal)
  fig <- ggbarplot(df, x = "method", y = "count", add = "mean_se", fill = "ifcausal", palette = colors, legend = "none", ...) + grids(linetype = "dashed")
  fig
}

f1 <- plot_ncausal_FDP(configtag, runtag,  simutaglist[[1]], colors = colset, cut = 0.2, ylim= c(0,225), main = "high gene PVE")
f2 <- plot_ncausal_FDP(configtag, runtag,  simutaglist[[2]], colors = colset, cut = 0.2, ylim= c(0,225), main = "low gene PVE")
gridExtra::grid.arrange(f1, f2, ncol=2)

f1 <- plot_ncausal_FDP(configtag, runtag,  simutaglist[[1]], colors = colset, cut = 0.4, ylim= c(0,225), main = "high gene PVE")
f2 <- plot_ncausal_FDP(configtag, runtag,  simutaglist[[2]], colors = colset, cut = 0.4, ylim= c(0,225), main = "low gene PVE")
gridExtra::grid.arrange(f1, f2, ncol=2)

f1 <- plot_ncausal_FDP(configtag, runtag,  simutaglist[[1]], colors = colset, cut = 1, ylim= c(0,225), main = "high gene PVE")
f2 <- plot_ncausal_FDP(configtag, runtag,  simutaglist[[2]], colors = colset, cut = 1, ylim= c(0,225), main = "low gene PVE")
gridExtra::grid.arrange(f1, f2, ncol=2)
```

@@

  if (isTRUE(rerun_ctwas)){
      ld_exprfs <- paste0(outputdir, runtag, "_simu", simutag, "_chr", 1:22, ".expr.gz")
      load(paste0(outputdir, runtag, "_simu", simutag, "_config", configtag, ".s2.susieIrssres.Rd"))
      source(paste0(outputdir, "config", configtag, ".R"))
      group_prior <- group_prior_rec[, ncol(group_prior_rec)] 
      group_prior[2] <- group_prior[2] * thin
      group_prior_var <- group_prior_var_rec[, ncol(group_prior_var_rec)]
      temp_reg <- data.frame("chr" = paste0("chr",chr), "start" = startpos, "stop" = endpos)
      write.table(temp_reg, file= "temp_reg.txt" , row.names=F, col.names=T, sep="\t", quote = F)
      z_gene <-  a[a$type == "gene", c("id", "t.value")]
      colnames(z_gene) <- c("id", "z")
      z_snp <-  a[a$type == "SNP", c("id", "alt", "ref", "t.value")]
      colnames(z_snp) <- c("id", "A1", "A2", "z")
      ctwas_rss(z_gene, z_snp, ld_exprfs= ld_exprfs, ld_pgenfs = ld_pgenfs, ld_R_dir = NULL, ld_regions_custom = "temp_reg.txt", thin = 1, outputdir = ".", outname = "temp", ncore = 1, ncore.rerun = 1, prob_single = 0,  group_prior = group_prior, group_prior_var = group_prior_var, estimate_group_prior = F, estimate_group_prior_var = F)
      a2 <- fread("temp.susieIrss.txt", header = T)
      a <- merge(a2, b, by = "id", all = T)
      # file.remove("temp.susieIrss.txt")
      # file.remove("temp.temp.susieIrssres.Rd")
      # file.remove("temp.regions.txt")
      # file.remove("temp_reg.txt")
   }

@@

### coloc false positives
```{r regionplot4, echo=FALSE,results='hide',fig.keep='all', message=F}
simutag <- "3-4"
a <- plot_region(runtag, simutag, configtag, chr = 1, startpos = 64000000 , endpos =66000000, rerun_ctwas = F, plot_eqtl = T )
```
```{r regionplot4-1, echo=FALSE,results='hide',fig.keep='all', message=F}
simutag <- "3-4"
a <- plot_region(runtag, simutag, configtag, chr = 1, startpos = 66000000 , endpos =70000000, rerun_ctwas = F, plot_eqtl = T )
```
```{r regionplot4-2, echo=FALSE,results='hide',fig.keep='all', message=F}
simutag <- "3-4"
a <- plot_region(runtag, simutag, configtag, chr = 1, startpos = 155000000 , endpos =170000000, rerun_ctwas = F, plot_eqtl = T )
```
```{r regionplot4-3, echo=FALSE,results='hide',fig.keep='all', message=F}
simutag <- "3-4"
a <- plot_region(runtag, simutag, configtag, chr = 5, startpos = 47000000 , endpos =50000000, rerun_ctwas = F, plot_eqtl = T )
```

### 2023-05-08 analysis/index.Rmd
simulation paper figures
110k simulation

@@

>>>>

* [**Using UKbiobank data (45k samples), fusion weights Adipose, more settings with low SNP PVE **](simulation-ctwas-ukbWG-gtex.adipose_s80.45_03222023.html)

* [**Using UKbiobank data (113k samples), fusion weights Adipose, more settings with low SNP PVE **](simulation-ctwas-ukbWG-gtex.adipose_s200.113_03212023.html)

@@

## Paper

* [**Simulation figures**](Paper_figures_simulation.html)

### 2023-05-08 analysis/simulation-ctwas-ukbWG-gtex.adipose_s200.113_03212023.Rmd
110k simulation

---
title: "Simulations to test ctwas summary stats version, 45k samples, a few new setings required"
output: html_document
---

@@

```{r files}
pgenfn = "/home/simingz/causalTWAS/ukbiobank/ukb_pgen_s80.45/ukb-s80.45_pgenfs.txt"
ld_pgenfn = "/home/simingz/causalTWAS/ukbiobank/ukb_pgen_s80.45/ukb-s80.45.2_pgenfs.txt"
outputdir = "/home/simingz/causalTWAS/simulations/simulation_ctwas_rss_20230322/" # /
comparedir = "/home/simingz/causalTWAS/simulations/simulation_ctwas_rss_20230322_compare/"
runtag = "ukb-s80.45-adi"
configtags = 1
simutags = paste(rep(1:9, each = length(1:5)), 1:5, sep = "-")

@@

```{r parameters}
n.ori <- 80000 # number of samples
n <- pgenlibr::GetRawSampleCt(pgens[[1]])
p <- sum(unlist(lapply(pgens, pgenlibr::GetVariantCt))) # number of SNPs
J <- 8021 # number of genes
```

@@

plot_PIP <- function(configtag, runtag,  simutags){
   phenofs <- paste0(outputdir, "ukb-s80.45-adi", "_simu", simutags, "-pheno.Rd")
   susieIfs <- paste0(outputdir, runtag, "_simu",simutags, "_config", configtag,".susieIrss.txt")

@@

configtag <- 1
runtag = "ukb-s80.45-adi"

simutags <- paste(1, c(1:5), sep = "-")
plot_par(configtag, runtag, simutags)
plot_PIP(configtag, runtag, simutags)
# plot_fusion_coloc(configtag, runtag, simutags)
# # plot_focus(configtag, runtag, simutags)
# print("FOCUS result is null")
# plot_smr(configtag, runtag, simutags)

simutags <- paste(2, 1:5, sep = "-")
plot_par(configtag, runtag, simutags)
plot_PIP(configtag, runtag, simutags)
# plot_fusion_coloc(configtag, runtag, simutags)
# #plot_focus(configtag, runtag, simutags)
# plot_smr(configtag, runtag, simutags)

simutags <- paste(3, 1:5, sep = "-")
plot_par(configtag, runtag, simutags)
plot_PIP(configtag, runtag, simutags)
# plot_fusion_coloc(configtag, runtag, simutags)
# #plot_focus(configtag, runtag, simutags)
# plot_smr(configtag, runtag, simutags)

simutags <- paste(4, 1:5, sep = "-")
plot_par(configtag, runtag, simutags)
plot_PIP(configtag, runtag, simutags)
# plot_fusion_coloc(configtag, runtag, simutags)
# #plot_focus(configtag, runtag, simutags)
# plot_smr(configtag, runtag, simutags)

simutags <- paste(5, c(1:5), sep = "-")
plot_par(configtag, runtag, simutags)
plot_PIP(configtag, runtag, simutags)
# plot_fusion_coloc(configtag, runtag, simutags)
# #plot_focus(configtag, runtag, simutags)
# plot_smr(configtag, runtag, simutags)

simutags <- paste(6, 1:5, sep = "-")
plot_par(configtag, runtag, simutags)
plot_PIP(configtag, runtag, simutags)
# plot_fusion_coloc(configtag, runtag, simutags)
# #plot_focus(configtag, runtag, simutags)
# plot_smr(configtag, runtag, simutags)

simutags <- paste(7, 1:5, sep = "-")
plot_par(configtag, runtag, simutags)
plot_PIP(configtag, runtag, simutags)
# plot_fusion_coloc(configtag, runtag, simutags)
# #plot_focus(configtag, runtag, simutags)
# plot_smr(configtag, runtag, simutags)

simutags <- paste(8, 1:5, sep = "-")
plot_par(configtag, runtag, simutags)
plot_PIP(configtag, runtag, simutags)
# plot_fusion_coloc(configtag, runtag, simutags)
# #plot_focus(configtag, runtag, simutags)
# plot_smr(configtag, runtag, simutags)

simutags <- paste(9, 1:5, sep = "-")
plot_par(configtag, runtag, simutags)
plot_PIP(configtag, runtag, simutags)
# plot_fusion_coloc(configtag, runtag, simutags)
# #plot_focus(configtag, runtag, simutags)
# plot_smr(configtag, runtag, simutags)
```

>>>>

---
title: "Simulations to test ctwas summary stats version, 113k samples, a few new setings required"
output: html_document
---

@@

```{r files}
pgenfn = "/home/simingz/causalTWAS/ukbiobank/ukb_pgen_s200.113/ukb-s200.113_pgenfs.txt"
ld_pgenfn = "/home/simingz/causalTWAS/ukbiobank/ukb_pgen_s200.113/ukb-s200.113.2_pgenfs.txt"
outputdir = "/home/simingz/causalTWAS/simulations/simulation_ctwas_rss_20230321/" # /
comparedir = "/home/simingz/causalTWAS/simulations/simulation_ctwas_rss_20230321_compare/"
runtag = "ukb-s200.113-adi"
configtags = 1
simutags = paste(rep(1:9, each = length(1:5)), 1:5, sep = "-")

@@

```{r parameters}
n.ori <- 200000 # number of samples
n <- pgenlibr::GetRawSampleCt(pgens[[1]])
p <- sum(unlist(lapply(pgens, pgenlibr::GetVariantCt))) # number of SNPs
J <- 8021 # number of genes
```

@@

plot_PIP <- function(configtag, runtag,  simutags){
   phenofs <- paste0(outputdir, runtag, "_simu", simutags, "-pheno.Rd")
   susieIfs <- paste0(outputdir, runtag, "_simu",simutags, "_config", configtag,".s1.susieIrss.txt")

@@

configtag <- 1
runtag = "ukb-s200.113-adi"

simutags <- paste(1, c(1:24,26:54, 56:70), sep = "-")
plot_par(configtag, runtag, simutags)
plot_PIP(configtag, runtag, simutags)
#plot_fusion_coloc(configtag, runtag, simutags)
# plot_focus(configtag, runtag, simutags)
#print("FOCUS result is null")
#plot_smr(configtag, runtag, simutags)

simutags <- paste(2, 1:30, sep = "-")
plot_par(configtag, runtag, simutags)
plot_PIP(configtag, runtag, simutags)
# plot_fusion_coloc(configtag, runtag, simutags)
# #plot_focus(configtag, runtag, simutags)
# plot_smr(configtag, runtag, simutags)

simutags <- paste(3, 1:10, sep = "-")
plot_par(configtag, runtag, simutags)
plot_PIP(configtag, runtag, simutags)
# plot_fusion_coloc(configtag, runtag, simutags)
# #plot_focus(configtag, runtag, simutags)
# plot_smr(configtag, runtag, simutags)

simutags <- paste(4, 1:10, sep = "-")
plot_par(configtag, runtag, simutags)
plot_PIP(configtag, runtag, simutags)
# plot_fusion_coloc(configtag, runtag, simutags)
# #plot_focus(configtag, runtag, simutags)
# plot_smr(configtag, runtag, simutags)

simutags <- paste(5, c(1:10), sep = "-")
plot_par(configtag, runtag, simutags)
plot_PIP(configtag, runtag, simutags)
# plot_fusion_coloc(configtag, runtag, simutags)
# #plot_focus(configtag, runtag, simutags)
# plot_smr(configtag, runtag, simutags)
```
